---
title: "Stratified EEG Connectivity Analysis"
output: pdf_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rstatix)
library(car)
library(vegan)

base_dir <- "D:/EEG/R-EEG"

methods <- c("dSPM", "eLORETA")
conn_types <- c("imcoh", "wpli2_debiased")
bands <- c("theta", "alpha", "beta")
```

### load demographics & connectivity data

```{r}
demo <- read_csv(file.path(base_dir, "Participants_LEMON.csv")) %>%
  rename(subject = `...1`) %>%
  mutate(
    subject = str_remove(subject, "^sub-"),
    subject = str_trim(subject),
    
    age_low = as.numeric(str_extract(Age, "^\\d+")),
    age_high = as.numeric(str_extract(Age, "\\d+$")),
    age_mid = (age_low + age_high) / 2,
    
    age_group_main = if_else(age_mid <= 60, "young", "old"),
    
    age_group_ablation = case_when(age_mid < 35 ~ "young", age_mid > 60 ~ "old",
      TRUE ~ NA_character_))

read_conn <- function(inv_method) {
  path <- file.path(
    base_dir,
    paste0("connectivity_", inv_method),
    "stats",
    "bootstrap_connectivity_summary.tsv"
  )
  
  read_tsv(path) %>%
    mutate(
      inverse_method = inv_method,
      subject = str_remove(subject, "^sub-"),
      subject = str_trim(subject)
    )
}

conn <- bind_rows(read_conn("dSPM"), read_conn("eLORETA"))

conn2 <- conn %>% left_join(demo, by = "subject")

conn2 %>% count(age_group_main)
conn2 %>% count(age_group_ablation)

demo %>%
  filter(age_mid >= 35, age_mid <= 60) %>%
  select(subject, Age, age_mid)

conn %>%
  semi_join(
    demo %>%
      filter(age_mid >= 35, age_mid <= 60), by = "subject") %>%
  count(subject)
```

### Stratified Wilcoxon age tests

```{r}
age_tests_main_stratified <- conn2 %>%
  filter(
    band %in% bands,
    method %in% conn_types,
    !is.na(age_group_main)
  ) %>%
  group_by(inverse_method, method, band, task) %>%
  summarise(
    n_young = sum(age_group_main == "young"),
    n_old = sum(age_group_main == "old"),
    mean_young = mean(bootstrap_mean[age_group_main == "young"], na.rm = TRUE),
    mean_old = mean(bootstrap_mean[age_group_main == "old"], na.rm = TRUE),
    p = ifelse(
      n_young > 0 & n_old > 0,
      wilcox.test(bootstrap_mean ~ age_group_main, exact = FALSE)$p.value,
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  group_by(inverse_method, method) %>%
  mutate(p_fdr_stratified = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  arrange(p_fdr_stratified)

age_tests_main_stratified

age_tests_ablation_stratified <- conn2 %>%
  filter(
    band %in% bands,
    method %in% conn_types,
    !is.na(age_group_ablation)
  ) %>%
  group_by(inverse_method, method, band, task) %>%
  summarise(
    n_young = sum(age_group_ablation == "young"),
    n_old = sum(age_group_ablation == "old"),
    mean_young = mean(bootstrap_mean[age_group_ablation == "young"], na.rm = TRUE),
    mean_old = mean(bootstrap_mean[age_group_ablation == "old"], na.rm = TRUE),
    p = ifelse(
      n_young > 0 & n_old > 0,
      wilcox.test(bootstrap_mean ~ age_group_ablation, exact = FALSE)$p.value,
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  group_by(inverse_method, method) %>%
  mutate(p_fdr_stratified = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  arrange(p_fdr_stratified)

age_tests_ablation_stratified
```

The demographic file contained five subjects between 35 and 60, but none of those subjects were present in the connectivity summary data. The main age split and the ablation split used the same effective connectivity sample.

### # ks tests

```{r}
ks_age_main <- conn2 %>%
  filter(band %in% bands, method %in% conn_types, !is.na(age_group_main)) %>%
  group_by(inverse_method, method, band, task) %>%
  summarise(
    n_young = sum(age_group_main == "young"),
    n_old = sum(age_group_main == "old"),
    ks_p = ifelse(
      n_young > 0 & n_old > 0,
      ks.test(
        bootstrap_mean[age_group_main == "young"],
        bootstrap_mean[age_group_main == "old"]
      )$p.value,
      NA_real_
    ),
    .groups = "drop") %>%
  group_by(inverse_method, method) %>%
  mutate(ks_p_fdr = p.adjust(ks_p, method = "BH")) %>%
  ungroup() %>%
  arrange(ks_p_fdr)

ks_age_main

ks_age_ablation <- conn2 %>%
  filter(band %in% bands, method %in% conn_types, !is.na(age_group_ablation)) %>%
  group_by(inverse_method, method, band, task) %>%
  summarise(
    n_young = sum(age_group_ablation == "young"),
    n_old = sum(age_group_ablation == "old"),
    ks_p = ifelse(
      n_young > 0 & n_old > 0,
      ks.test(
        bootstrap_mean[age_group_ablation == "young"],
        bootstrap_mean[age_group_ablation == "old"]
      )$p.value,
      NA_real_
    ),
    .groups = "drop") %>%
  group_by(inverse_method, method) %>%
  mutate(ks_p_fdr = p.adjust(ks_p, method = "BH")) %>%
  ungroup() %>%
  arrange(ks_p_fdr)

ks_age_ablation
```

#### box plot with jitter points

```{r}
ggplot(
  conn2 %>%
    filter(band == "alpha", method == "imcoh", !is.na(age_group_ablation)),
  aes(x = age_group_ablation, y = bootstrap_mean)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.15, alpha = 0.7) +
  facet_grid(inverse_method ~ task) + theme_bw() +
  labs(x = "Age Group", y = "Connectivity",
    title = "Alpha imcoh Connectivity: Young <35 vs Old >60"
  )
```

### ANOVA and ANCOVA

```{r}
anova_data <- conn2 %>%
  filter(
    band %in% bands,
    method %in% conn_types,
    !is.na(age_group_ablation)
  )

anova_model <- lm(
  bootstrap_mean ~ age_group_ablation * band * task * inverse_method * method,
  data = anova_data
)

Anova(anova_model, type = 3)

ancova_model <- lm(
  bootstrap_mean ~ age_mid * band * task * inverse_method * method,
  data = anova_data
)

Anova(ancova_model, type = 3)
```

### dSPM vs eLORETA using paired testing Wilcoxon-Signed rank test

```{r}
dspm_vs_eloreta <- conn2 %>%
  filter(
    band %in% bands,
    method %in% conn_types
  ) %>%
  select(subject, task, band, method, inverse_method, bootstrap_mean) %>%
  pivot_wider(
    names_from = inverse_method,
    values_from = bootstrap_mean
  ) %>%
  filter(!is.na(dSPM), !is.na(eLORETA)) %>%
  group_by(method, band, task) %>%
  summarise(
    mean_dSPM = mean(dSPM, na.rm = TRUE),
    mean_eLORETA = mean(eLORETA, na.rm = TRUE),
    p = wilcox.test(dSPM, eLORETA, paired = TRUE, exact = FALSE)$p.value,
    .groups = "drop"
  ) %>%
  group_by(method) %>%
  mutate(p_fdr_stratified = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  arrange(p_fdr_stratified)

dspm_vs_eloreta
```

### WPLI-debiased vs IMCOH using paired testing Wilcoxon-Signed rank test

```{r}
wpli_vs_imcoh <- conn2 %>%
  filter(
    band %in% bands,
    method %in% conn_types
  ) %>%
  select(subject, task, band, inverse_method, method, bootstrap_mean) %>%
  pivot_wider(
    names_from = method,
    values_from = bootstrap_mean
  ) %>%
  filter(!is.na(wpli2_debiased), !is.na(imcoh)) %>%
  group_by(inverse_method, band, task) %>%
  summarise(
    mean_wpli = mean(wpli2_debiased, na.rm = TRUE),
    mean_imcoh = mean(imcoh, na.rm = TRUE),
    p = wilcox.test(wpli2_debiased, imcoh, paired = TRUE, exact = FALSE)$p.value,
    .groups = "drop"
  ) %>%
  group_by(inverse_method) %>%
  mutate(p_fdr_stratified = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  arrange(p_fdr_stratified)

wpli_vs_imcoh
```

### PERMANOVA and PERMDISP
```{r}
run_permanova <- function(data, age_var, inv_method) {
  
  wide_data <- data %>%
    filter(inverse_method == inv_method, band %in% bands, method %in% conn_types, 
           !is.na(.data[[age_var]])) %>%
    select(subject, age_group = all_of(age_var), band, task, method, 
           bootstrap_mean) %>%
    unite(feature, band, task, method, sep = "_") %>%
    pivot_wider(names_from = feature, values_from = bootstrap_mean) %>%
    drop_na()
  
  print(wide_data %>% count(age_group))
  
  X <- wide_data %>% select(-subject, -age_group)
  
  meta <- wide_data %>% select(subject, age_group)
  
  permanova <- adonis2(X ~ age_group, data = meta, method = "euclidean", 
                       permutations = 9999)
  
  dist_mat <- dist(X)
  
  disp <- betadisper(dist_mat, meta$age_group)
  
  permdisp <- permutest(disp, permutations = 9999)
  
  list(data_counts = wide_data %>% count(age_group), permanova = permanova, 
       permdisp = permdisp)
  }

permanova_dspm_main <- run_permanova(conn2, age_var = "age_group_main", 
                                     inv_method = "dSPM")

permanova_eloreta_main <- run_permanova(conn2, age_var = "age_group_main",
                                        inv_method = "eLORETA")

permanova_dspm_ablation <- run_permanova(conn2, age_var = "age_group_ablation",
                                         inv_method = "dSPM")

permanova_eloreta_ablation <- run_permanova(conn2, age_var = "age_group_ablation", 
                                            inv_method = "eLORETA")

permanova_dspm_main$permanova
permanova_dspm_main$permdisp

permanova_eloreta_main$permanova
permanova_eloreta_main$permdisp

permanova_dspm_ablation$permanova
permanova_dspm_ablation$permdisp

permanova_eloreta_ablation$permanova
permanova_eloreta_ablation$permdisp
```